---
title: "Reproducible Research-Week 2 Project"
author: "Yvonne"
date: "November 7, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(plyr)
library(dplyr)
library(lattice) 
```

The following codes will:

A)  Loading and preprocessing the data
    1)  It reads the activity.csv file into workspace
    2)  Remove cases with NA
    
```{r activity}
activity <- read.csv("activity.csv")        # read csv file
noNAData <- activity[complete.cases(activity), ]  #drop na 
```

B) What is mean total number of steps taken per day?
    1) compute the total steps made per day
    2) plot a histogram and boxplot for the total number of steps taken each day
    3) Calculate and report the mean and median of the total number of steps taken per day

```{r total_steps_perday}

#compute total of steps: 570608
sum(activity$steps,na.rm=TRUE)        

# draw boxplot
qplot(date, steps, data = activity, geom = "boxplot")+xlab("Day") + ylab("Steps")+ggtitle("Histogram of Total Number of Steps Per Day")+ theme(axis.text.x = element_text(angle = 90, hjust = 1))

# draw histogram
ggplot(activity, aes(as.factor(date),steps))+geom_bar(fill="purple", stat="identity")+xlab("Day") + ylab("Steps")+ggtitle("Histogram of Total Number of Steps Per Day")+ theme(axis.text.x = element_text(angle = 90, hjust = 1))

# mean no of steps each day: 10766
as.integer(mean(aggregate(steps~date,activity, FUN=sum)$steps))

# mean no of steps each day: 10765
as.integer(median(aggregate(steps~date,activity, FUN=sum)$steps))

```

C) What is the average daily activity pattern?
```{r daily_activity}
Q2 <- aggregate(steps~interval, noNAData, FUN=mean)
plot(Q2$steps ~ Q2$interval, type="l", xlab="Interval", ylab="Average Steps per Day (5-mins interval")
title(main = "Average Steps Per 5-min Interval")
# Q2[Q2$steps==max(Q2$steps),]  or use which.max function
Q2[which.max(Q2$steps),]  # interval when steps is max: 835

```

D) Imputing missing values

```{r missing_value}
#1. the total number of rows with NAs
sum(is.na(activity))  #2304
#2. fil in the missing values with mean for that day
#3. Create a new dataset that equal to  original dataset but with missing data filled
fillNA <- activity  #new dataset
meanInt<- tapply(fillNA$steps, fillNA$interval, mean, na.rm=TRUE, simplify = TRUE)
nas<- is.na(fillNA$steps)#get position of steps = na
fillNA$steps[nas] <- meanInt[as.character(fillNA$interval[nas])]

#4. Draw histogram using dataset without NA records
ggplot(fillNA, aes(as.factor(date),steps))+geom_bar(fill="purple", stat="identity")+xlab("Day") + ylab("Steps")+ggtitle("Histogram of Total Number of Steps Per Day (filled NA)")+ theme(axis.text.x = element_text(angle = 90, hjust = 1))

# mean of steps taken per day: 10766
as.integer(mean(aggregate(steps~date, fillNA, FUN=sum)$steps))

# median of steps taken per day: 10766
as.integer(median(aggregate(steps~date, fillNA, FUN=sum)$steps))

# The new mean of the imputed data is 10766 steps compared to the old mean of 10766 steps. There is no difference on average per day.
# Same explanation for mean

```

E) Are there differences in activity patterns between weekdays and weekends?

``` {r weekdays_weekends}
#1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" 
fillNA<- fillNA%>%
        mutate(weektype= ifelse(weekdays(as.Date(fillNA$date))=="Saturday" | weekdays(as.Date(fillNA$date))=="Sunday", "Weekend", "Weekday"))

head(fillNA)

#2. Make a panel plot containing a time series plot 
## Summarize data by interval and type of day
dayInt <- ddply(fillNA, .(interval, weektype), summarize, Avg = mean(steps))
head(dayInt)

xyplot(Avg~interval|weektype, data=dayInt, type="l",  layout = c(1,2),
       main="Average Steps per Interval Based on Type of Day", 
       ylab="Average Number of Steps", xlab="Interval")

# Answer: there differences in activity patterns between weekdays and weekends.  Weekday # has higher spike in the earlier part of the day. But Weekday has more even spike over
# the day.
```